Cover Times of Random Walks

### 2019-05-14 paper.tex
final proofread

\abstract{
The cover time of a graph is the number of steps a 
discrete random walk takes to visit vertex at least once.
We begin by investigating the cover time of small and structured graphs but
quickly find that our precise strategy fails
for larger, unstructured ones.
To bound the cover time on arbitrary graphs,
we apply more general probablistic arguments and an important electrical
network identity.
Finally, we describe the limiting distribution of the cover time of
complete graphs with self-edges and apply our analysis to an intuitive example.
}

@@

Next consider a particular case of coupon collecting called the birthday problem.
Assuming all 365 days of the year, excluding February 29, are equally likely,
how many people would we need to ensure each birthday is represented
by at least one person?
A lower bound is certainly 365 but it is possible to get an arbitrarily
large population without covering each birthday.

@@

Define the weight of vertex $i$ 
\begin{align}
C_i &= \sum_{j:i \sim j} C_{i,j}. \nonumber
\end{align}
In the unweighted case, $C_i$ is simply the degree of vertex $i$.

@@

Let $T_{j}$ be the number of steps until the first visit to $j$
after time time $t=0$.
Formally,
\begin{align}
T_{j} &= \textrm{min} \left\{t > 0 : X_t = j \right\}. \nonumber
\end{align}
Let $\E_i[T_j] = \E[T_j | X_0 = i]$.
This is the expected time of the first visit to $j$ given
that the random walk started in vertex $i$.
Then $\E_j[T_j]$ is the return time to $j$.
Now define the hitting time $\hit$ for the graph
be the maximum expected time between two vertices on $G$.

Formally, 
\begin{align}
\hit &= \max_{i,j \in V(G)} \E_i[T_j]. \nonumber
\end{align}
Let the cover time variable $\cov$ be the first time all vertices have been
visited by the random walk $(X_t)$.
Formally, $\cov$ is the minimum time that for all $i \in V(G)$
there exists $t \leq \cov$ such that $X_t = i$.
We define the cover time as the expected value of $\cov$ from
the worst initial vertex.
Formally,
\begin{align}
\tcov &= \max_{j \in V(G)} \E_j[\cov] . \nonumber
\end{align}

@@

So that the reader may more easily distinguish between
vertices, consider $x$ and $y$ in $V(G)$.
Recall that $C_{x,y}$ is the weight of the edge
between vertices $x$ and $y$, $C_x$ is the sum
of all $C_{x,y}$ where $x \sim y$, and the probability $\pr_{x,y}$
that we go from $x$ to $y$ is $C_{x,y}/C_x$.
For the rest of this section, we will call $C_{x,y}$
the conductance between $x$ and $y$.
Define the resistance $R_{x,y}$ between $x$ and $y$
as $1/ C_{x,y}$.

@@

We verify that $\pi$ is the stationary distribution
of $(X_t)$.
To begin,
\begin{align}
C_x \pr_{x,y} &= C_x \frac{C_{x,y}}{C_x} = C_{x,y} = C_{y,x} 
= C_y \frac{C_{y,x}}{C_y} = C_y \pr_{y,x}. \nonumber
\end{align}
Dividing by $C$ gives $\pi_x \pr_{x,y} = \pi_y \pr_{y,x}$.
(So $\pi$ satisfies detailed balance.)
Summing over $x$,
\begin{align}
\sum_x \pi_x \pr_{x,y} = \pi_y . \nonumber
\end{align}
Then $\pi \pr = \pi$ where $\pr$ is the matrix of transition probabilities.
It follows that $\pi$ is indeed the stationary distribution of $(X_t)$.

We now introduce voltage, current, and effective resistance.
Consider $a$ and $b$ in $V(G)$.
Set the voltage to be 1 at $a$ and 0 at $b$.
We may think of voltage as the probability we get to $a$ before $b$.
Probabilistically, of course $v_a = 1$ and $v_b=0$.
The voltage at every vertex $x$ in  $V(G) \textbackslash \{a,b\}$
is determined by the boundary values (at $a$ and $b$).

The current between two vertices is the product of the
difference in their voltages and the conductance of their edge.
We may think of current as the net number of movements through an edge.
Kirchoff's Current Law says that the sum of current on any vertex is 0.
Probabilistically, of course every time we enter a vertex we must also leave it.
Using Kirchoff's Current Law and the definition of current,
we justify the probabilistic interpretation of voltage
\begin{align}
\sum_y i_{x,y} &= 0 = \sum_y (v_x - v_y)C_{x,y} \nonumber \\
v_x \sum_y C_{x,y} &= \sum_y C_{x,y} v_y \nonumber \\
v_x &= \sum_y \frac{C_{x,y}}{C_x} v_y = \sum_y \pr_{x,y} v_y. \nonumber
\end{align}
We can verify that $v_x = \sum_y \pr_{x,y} v_y$
using first step analysis on the probability of reaching $a$ before $b$.

@@

\cref{sec:bounds} deals with the large or unstructured graphs 
for which we must use bounds.
We begin by establishing an important identity between
commute times and effective resistance in \cref{sec:commute}.
(Our electrical network analysis is an extension of \cite{DS84}
and we use \cite{LP17} for the next two subsections.)
We take our strategy for finding the cover time of the complete graph
from \cref{sec:complete} and apply it to general graphs in \cref{sec:matthews}.
In \cref{sec:binary}, we use both the commute time identity and the Matthews Method
to bound the cover time of a binary tree.
In \cref{sec:spanning}, we follow \cite{AF14} and use the idea of a spanning tree
as well as the commute time identity to develop another bound for general graphs.

Up until \cref{sec:dist}, we have only considered the expected cover time.
We now delve into the distribution of the cover time of a complete graph.
We draw on \cite{Du11} and find that the cover time variable approaches a Poisson
distribution.
In \cref{sec:birthday}, we apply \cref{sec:complete_dist} to the birthday problem.
(The R code for our computations and \cref{fig:birthday} can be found in \cref{app:birthday}.)

@@

\cref{fig:example} shows our small graph and the first step analysis
for the random walk $(X_t)$ started from the top left vertex.
The black circle indicates the current vertex of the random walk
and the empty circles indicate the visited vertices.

@@

With a little linear algebra, we find that $\tcov = 43/5$.
We leave it as an exercise to the reader to verify that the time
to visit all vertices from the other starting vertex is $43/6$.

For large graphs, the first step analysis as an approach to finding the expected
cover time is obviously intractable.
The rest of this section deals with particularly symmetric graphs
for which it is possible to find a solution for an arbitrary $n$.
However, we must rely on cover time bounds for the vast majority of cases.

@@

\begin{figure}[ht]
	\centering
	\begin{tikzpicture}[scale=.1]
	  \graph[circular placement, radius=3cm,
	         empty nodes, nodes={circle,draw,fill=black,scale=.7}] {
	    \foreach \x in {a,...,f} {
	      \foreach \y in {\x,...,f} {
	        \x -- \y;
	      };
	    };
	  };
	  \foreach \x [count=\idx from 0] in {a,...,f} {
	    \pgfmathparse{90 + \idx * (360 / 6)}
	    \node at (\pgfmathresult:4.4cm) {};
	  };
	\end{tikzpicture}
    \caption{The complete graph with six vertices (and ten edges).}\label{fig:complete}	
\end{figure}

Our goal is to find the expected number of steps until we have visited all
$n$ vertices in the complete graph.
Observe that the random walk can start at any vertex without loss of generality
because the graph is symmetric.
The strategy is to write the cover time $\tcov$ in terms of the expected
value of more simple random variables.
Let $X_i$ be the random variable that represents the number of steps to go
from $i-1$ to $i$ unique vertices (excluding the current vertex).

@@

\begin{comment}
It follows that $X_i$ is geometric with distribution
\begin{align}
\pr(X_i = k) = (1-p_i)^{k-1}p_i, \nonumber
\end{align}
where $p_i$ is the probability of moving to a previously unvisited vertex.
\end{comment}
We now find the probability $p_i$ that 
we go to a previously unvisited vertex is $(n-i+1)/(n-1)$.
This is because there are $n-i+1$ unvisited 
vertices and a total of $n-1$ adjacent vertices.
It follows that $\E[X_i] = (n-1)/(n-i+1)$.
Then
\begin{align}
\tcov&=   \sum_{i=1}^{n} \frac{n-1}{n-i+1} \nonumber \\
&=  (n-1) \left(\frac{1}{n-1}+\frac{1}{n-2}+\cdots+1\right). \nonumber
\end{align}

A natural interpretation of the cover time on a complete graph is 
the coupon collector problem we introduced in \cref{sec:intro}.
There are $r$ distinct coupons one of which arrives in the mail everyday.
We want to know how many days until the collector has collected each coupon.
The strategy is to think of every coupon as a vertex on the complete graph.
We then move from vertex to vertex with uniform probability.
The difference between the coupon collector problem and the complete graph
is that we now have self-edges
whereas before we had to leave our current vertex at each step.
That is, we can stay in the same vertex (if the collector receives the same coupon
two days in a row).
We apply our approach to the complete graph and substitute the $r$ possible coupons
we can find for the $n-1$ vertices we could move to.
Then
\begin{align}
\tcov&=  r \left(\frac{1}{r}+\frac{1}{r-1}+\cdots+1\right). \nonumber
\end{align}

@@

We will begin by considering the time $(X_t)$ takes to
visit either one of the endpoints from vertex $i$.

From $X_t=i$, we move with equal probability to $i+1$
and $i-1$ until we reach either $0$ or $N$.

@@

where $\E_{i}[T_{i+1}]$ is the expected time from state $i$ to $i+1$.
Since there is only one edge incident to 0, we always move from 0 to 1.
From state $i\neq0$, we take one step in each direction with equal probability.
In half the cases, we have arrived at $i+1$.
In the other half of cases, we are in $i-1$.
It will take us $\E_{i-1}[T_i]$ steps back to $i$ and then another
$\E_i[T_{i+1}]$ steps to $i+1$.
Formally,
\begin{align}
\E_{i}[T_{i+1}] &= 1 + \frac{1}{2}(\E_{i-1}[T_i]+\E_i[T_{i+1}]) \nonumber \\
 &= 2 + \E_{i-1}[T_{i}]. \nonumber
\end{align}

@@

Then
\begin{align}
\E_0[T_N] &= 1 + 3 + \cdots + 2i - 1 + \cdots + 2(N-1) -1 \nonumber \\
&= N^2 \nonumber
\end{align}

Then, by \cref{thm:line}, for the random walk from state $i$ on the line graph,
\begin{align}
\tcov = i(N-i) + N^2. \nonumber
\end{align}

@@

Consider a random walk on a simple cycle as in \cref{fig:cycle}.
(Note that the starting vertex is arbitrary since the graph is symmetric.)
Let $(X_t)$ be the walk on the number line (between $-n$ and $n$)
where $X_0 = 0$.
The cover time $\tcov$ of the simple cycle is the expected number of steps 
until we visit $n$ distinct integers.

@@

\section{Bounds}\label{sec:bounds}

@@

Recall $\sum_y P_{a,y} v_y$ is the probability we reach $a$ before $b$
so 1 minus that is the probability we reach $b$ before returning to $a$.

@@

Recall that $\pi_a$ is the probability that $(X_t) = a$.
By a similar argument, $1/\pi_a$ is the expected return time to $a$.
If we condition on reaching $b$ before turning to $a$,
\begin{align}
\frac{1}{\pi_a} &= \E_a[T_a] =
\E_a[T_a|T_b < T_a] \pr(T_b<T_a) +
\E_a[T_a|T_a < T_b] \pr(T_a<T_b) \nonumber.
\end{align}

The expected commute time between $a$ and $b$ is the
expected return time to $a$ given we reach $b$ plus
the expected return time to $a$ given we do not reach $b$ times
the number of times we return to $a$ before the success when
we reach $b$.
Formally, we write
\begin{align}
\E_a[T_b] + \E_b[T_a] &= 
\E_a[T_a | T_b < T_a] + 
\E_a[T_a | T_a < T_b] (\frac{1}{\pr(T_b<T_a)} - 1) \nonumber.
\end{align}
Observe that $1/\pr(T_b<T_a)-1 = (1-\pr(T_b<T_a))/\pr(T_b<T_a)$, 
which simplifies to $\pr(T_a<T_b)/\pr(T_b<T_a)$.
Multiplying by the probability we reach $b$ before returning to $a$,
\begin{align}
\E_a[T_a|T_b < T_a] \pr(T_b<T_a) + \E_a[T_a|T_a<T_b] \pr(T_a<T_b)=\frac{1}{\pi_a} \nonumber
\end{align}
and the commute identity follows
\begin{align}
E_a[T_b] + E_b[T_a] = \frac{1}{\pi_a}\frac{1}{\pr_a(T_b < T_a)} = \Reff C.
\label{eqn:commute}
\end{align}

@@

Our goal is to bound the cover time in terms of the hitting time.

Consider the first term of the sum.
Recall that the hitting time $\hit$ is the maximum
expected time from any vertex to any other.
It follows that for any choice of $\sigma_1$,
the expected time from $n$ to $\sigma_1$ is less than or equal to $\hit$.
Formally for $s$ an arbitrary state such that $1\leq s \leq n-1$,
$\E_n[T_1] = \E_n[T_s] \leq \hit$.

@@

Since $\sigma$ is a random permutation, the probability that
we visit $\sigma_k$ last is $1/k$.
Formally, $P(L_k = \sigma_k) = 1/k$.

Finally, we find the expected time between $T_{k-1}$ and $T_{k}$
given that last vertex we visit is $\sigma_k$.
Consider $r$ and $s$ where $1 \leq r \neq s \leq n-1$
such that $L_{k-1} = r$ and $L_k = s$.
Then $\E_n[T_k - T_{k-1} | L(k) = \sigma_k] = \E_r[T_s] \leq \hit$.
So
\begin{align}
\E_n[T_k - T_{k-1}] &\leq
\hit \frac{1}{k}. \nonumber 
\end{align}

@@

The same strategy for the Matthews upper bound can be used
to find a lower bound.
Instead of looking for all $n$ vertices, we search
for $A \subseteq V(G)$.
The Matthews lower bound will always hold.
However, we can tighten the bound by picking $A$ such
that the hitting time is large between any pair of vertices.
We use large in a relative sense and 
offer an illustrative example in \cref{sec:binary}.

@@

\begin{proof}
Without loss of generality, label the vertices $\{1, 2, \ldots, n\}$.
We may assume that the walk started at vertex $x \in A$.
Let $\sigma$ be a uniform permutation of the vertices in $A$.
The strategy is to look for the vertices in order $\sigma$.

@@

Our goal is to bound the cover time in terms of $\tmina$.

Consider the first term of the sum.
With probability $1/|A|$, $\sigma_1 = x$ and $T_1=0$.
Otherwise, the expected time from $x$ to $T_1$ is less than or equal to
$\tmina$.
Then
\begin{align}
\E_x[T_1] \geq \frac{1}{|A|}0 + \frac{|A| -1}{|A|} \tmina
= \left(1 - \frac{1}{|A|} \right) \tmina \nonumber
\end{align}

@@

Since $\sigma$ is a random permutation, $P(L_k = \sigma_k) = 1/k$.

Now find the expected time between $T_{k-1}$ and $T_{k}$
given that last vertex we visit is $\sigma_k$.
Consider $r, s \in A$ such that $L_{k-1} = r$ and $L_k = s$.
Then $\E_x[T_k - T_{k-1} | L(k) = \sigma_k] = \E_r[T_s] \geq \tmina$.
So
\begin{align}
\E_x[T_k - T_{k-1}] &\geq
\tmina \frac{1}{k}. \nonumber 
\end{align}

@@

We now use the Matthews Lower Bound.
Choose a set $A \subseteq V(G)$ with a large hitting time between any
pair of vertices.
To find such a set, fix a level $h$ in our binary tree.
Then let $A$ be a set of $2^h$ leaves so that each vertex at level $h$
has a unique descendent in $A$.
As $h$ increases, the distance between the leaves in $A$ decreases.
Thus we want to choose a smaller value of $h$ so that our lower bound will be bigger
(and better).

@@

A spanning tree $\T$ is a subset of a graph with all edges
in $V(G)$ connected by the minimum number of edges in $E(G)$.
\cref{fig:span} is an example where the bold edges indicate the edges in $E(\T)$
and the light edges indicate the edges in $E(G)$ but not $E(\T)$.
For the connected graphs we consider, the minimum number of edges
needed to connect all vertices is $n-1$.
Formally, $V(\T) = V(G)$ and $E(\T) \subseteq E(G)$
such that $|E(\T)| = n-1$ and $\T$ is connected.

@@

The Poisson approximation of the binomial implies that if $n \rightarrow \infty$
and $r/n \rightarrow c$, then the number of visits to each vertex will approach
a Poisson distribution with mean $c$.
We can use this to intuite our conclusion: the proportion of visits to each vertex
approaches $e^{-c}$.

@@

Define $p_m(r,n)$ as the probability that exactly $m$ vertices are unvisited
when $r$ steps have been taken.
Then the probability that all vertices have been visited is the complement
of the probability that at least one vertex has not been visited.

@@

We bound the expression from the top and the bottom and use the squeeze
theorem to prove the most recent limit.
We will first show the upper bound.
Notice that $(1-x) \leq e^{-x}$ and recall our assumption that
$ne^{r/n} \rightarrow \lambda$.
Then
\begin{align}
{n \choose m} \left(1-\frac{m}{n}\right)^r &\leq \frac{n^m}{m!} e^{mr/n} 
= \frac{(n e^{r/n})^m} {m!} \nonumber \\
&\rightarrow \frac{\lambda ^m} {m!}. \nonumber
\end{align}

@@

\subsection{The Birthday Problem}\label{sec:birthday}
Consider a complete graph (with self-edges) and 365 vertices.
We may think of the cover time variable of this graph as the number of people
until every birthday is represented.
We call this the birthday problem.

>>>>

\abstract{
The cover time of a graph is the number of steps a 
discrete random walk takes to visit every vertex at least once.
We begin by investigating the cover time of small and structured graphs but
quickly find that our precise strategy fails
for larger, unstructured ones.
To bound the cover time on arbitrary graphs,
we apply more general probablistic arguments and an important electrical
network identity.
Finally, we describe the limiting distribution of the cover time of
complete graphs with self-edges and apply our analysis to an intuitive example.
}

@@

Next consider a particular case of coupon collecting called the birthday problem.
Assuming all 365 days of the year, excluding February 29, are equally likely,
how many people would we need to ensure each birthday is represented
by at least one person?
A lower bound is certainly 365 but it is possible to get an arbitrarily
large population without `covering' every birthday.

@@

Define the weight of vertex $i$ as
\begin{align}
C_i &= \sum_{j:i \sim j} C_{i,j}. \nonumber
\end{align}
In the unweighted case, $C_i$ is simply the degree of vertex $i$.

@@

Let $T_{j}$ be the number of steps until the first visit to $j$
after time time $t=0$.
Formally,
\begin{align}
T_{j} &= \textrm{min} \left\{t > 0 : X_t = j \right\}. \nonumber
\end{align}
Let $\E_i[T_j] = \E[T_j | X_0 = i]$.
This is the expected time of the first visit to $j$ given
that the random walk started in vertex $i$.
Then $\E_j[T_j]$ is the return time to $j$
since $t$ is strictly positive.
Now define the hitting time $\hit$ for the graph
be the maximum expected time between two vertices on $G$.
Formally, 
\begin{align}
\hit &= \max_{i,j \in V(G)} \E_i[T_j]. \nonumber
\end{align}
Let the cover time variable $\cov$ be the first time all vertices have been
visited by the random walk $(X_t)$.
Formally, $\cov$ is the minimum time that for all $i \in V(G)$
there exists $t \leq \cov$ such that $X_t = i$.
We define the cover time as the expected value of $\cov$ from
the worst initial vertex.
Formally,
\begin{align}
\tcov &= \max_{j \in V(G)} \E_j[\cov] . \nonumber
\end{align}

@@

So that the reader may more easily distinguish between
vertices, consider $x$ and $y$ in $V(G)$.
Recall that $C_{x,y}$ is the weight of the edge
between vertices $x$ and $y$, $C_x$ is the sum
of all $C_{x,y}$ where $x \sim y$, and the probability $\pr_{x,y}$
that we go from $x$ to $y$ is $C_{x,y}/C_x$.
For the rest of this section, we will call $C_{x,y}$
the conductance between $x$ and $y$.
Define the resistance $R_{x,y}$ between $x$ and $y$
as $1/ C_{x,y}$.
(We will soon justify the electrical network interpretation of
a random walk using Ohm's Law.)

@@

We verify that $\pi$ is the stationary distribution
of $(X_t)$.
To begin,
\begin{align}
C_x \pr_{x,y} &= C_x \frac{C_{x,y}}{C_x} = C_{x,y} = C_{y,x} 
= C_y \frac{C_{y,x}}{C_y} = C_y \pr_{y,x}. \nonumber
\end{align}
Dividing by $C$ gives $\pi_x \pr_{x,y} = \pi_y \pr_{y,x}$.
(So $\pi$ satisfies detailed balance.)
Summing over $x$,
\begin{align}
\sum_x \pi_x \pr_{x,y} = \pi_y . \nonumber
\end{align}
In the language of Markov Chains,
$\pi \pr = \pi$ where $\pr$ is the matrix of transition probabilities.
It follows that $\pi$ is the stationary distribution of $(X_t)$.
The takeaway for the reader unfamiliar with Markov Chains is that
$\pi_y$ is indeed the long term proportion of time spent in vertex $y$.

We now introduce voltage, current, and effective resistance.
Consider $a$ and $b$ in $V(G)$.
Set the voltage to be 1 at $a$ and 0 at $b$.
We may think of voltage as the probability we get to $a$ before $b$:
$v_a$ the probability we get to $a$ before $b$ from $a$
is 1 and $v_b$ the probability we get to $a$ before $b$ from $b$ is 0.
The voltage at every vertex $x$ in  $V(G) \textbackslash \{a,b\}$
is determined by the boundary values (at $a$ and $b$).

By Ohm's Law, the current between two vertices is the product of the
difference in their voltages and the conductance of their edge.
We may think of current as the net number of movements through an edge.
Kirchoff's Current Law says that the sum of current on any vertex is 0.
Probabilistically: every time we enter a vertex we must also leave it.
Using Kirchoff's Current Law and the definition of current,
we justify the probabilistic interpretation of voltage
\begin{align}
\sum_y i_{x,y} &= 0 = \sum_y (v_x - v_y)C_{x,y} \nonumber \\
v_x \sum_y C_{x,y} &= \sum_y C_{x,y} v_y \nonumber \\
v_x &= \sum_y \frac{C_{x,y}}{C_x} v_y = \sum_y \pr_{x,y} v_y. \nonumber
\end{align}
We can verify that $v_x = \sum_y \pr_{x,y} v_y$
using total probability on the likelihood of reaching $a$ before $b$ from $x$.

@@

\cref{sec:bounds} deals with the large or unstructured graphs 
for which we must use cover time bounds.
We begin by establishing an important identity between
commute times and effective resistance in \cref{sec:commute}.
(Our electrical network analysis is an extension of \cite{DS84}
and we use \cite{LP17} for the next two subsections.)
We take our strategy for finding the cover time of the complete graph
from \cref{sec:complete} and apply it to general graphs in \cref{sec:matthews}.
In \cref{sec:binary}, we use both the commute time identity and the Matthews Method
to bound the cover time of a binary tree.
In \cref{sec:spanning}, we follow \cite{AF14} and use the idea of a spanning tree
as well as the commute time identity to develop another bound for general graphs.

Up until \cref{sec:dist}, we only consider expected cover time.
We now delve into the distribution of the cover time on a complete graph with self-edges.
We draw on \cite{Du11} and find that the cover time variable approaches a Poisson
distribution.
In \cref{sec:birthday}, we apply \cref{sec:complete_dist} to the birthday problem.
(The R code for our computations and \cref{fig:birthday} can be found in \cref{app:birthday}.)

@@

\cref{fig:example} shows our small graph and the first step analysis
for the random walk $(X_t)$ started from the top right vertex.
The black circle indicates the current vertex of the random walk
and the empty circles indicate the visited vertices.

@@

With a little linear algebra, we find that $\tcov = 43/5$.
We leave it as an exercise to the reader to verify that the time
to visit all vertices from the other (i.e. symmetrically distinct) vertex is $43/6$.

For large graphs, the first step analysis as an approach to finding the expected
cover time is obviously intractable.
The rest of this section deals with particularly symmetric graphs
for which it is possible to find a solution for an arbitrary number of vertices $n$.
However, we must rely on cover time bounds for the vast majority of cases.

@@

\begin{figure}[ht]
	\centering
	\begin{tikzpicture}[scale=.1]
	  \graph[circular placement, radius=3cm,
	         empty nodes, nodes={circle,draw,fill=black,scale=.7}] {
	    \foreach \x in {a,...,f} {
	      \foreach \y in {\x,...,f} {
	        \x -- \y;
	      };
	    };
	  };
	  \foreach \x [count=\idx from 0] in {a,...,f} {
	    \pgfmathparse{90 + \idx * (360 / 6)}
	    \node at (\pgfmathresult:4.4cm) {};
	  };
	\end{tikzpicture}
    \caption{The complete graph with six vertices (and 15 edges).}\label{fig:complete}	
\end{figure}

Our goal is to find the expected number of steps until we have visited all
$n$ vertices in the complete graph.
Observe that the random walk can start at any vertex without loss of generality
because the graph is symmetric.
The strategy is to write the cover time $\tcov$ in terms of the expected
value of more simple random variables.
Let $X_i$ be the random variable that represents the number of steps to go
from $i-1$ to $i$ distinct vertices.

@@

\begin{comment}
It follows that $X_i$ is geometric with distribution
\begin{align}
\pr(X_i = k) = (1-p_i)^{k-1}p_i, \nonumber
\end{align}
where $p_i$ is the probability of moving to a previously unvisited vertex.
\end{comment}
We now find the probability $p_i$ that 
we go to a previously unvisited vertex is $(n-i+1)/(n-1)$.
This is because there are $n-i+1$ unvisited 
vertices and a total of $n-1$ adjacent vertices.
It follows that $\E[X_i] = (n-1)/(n-i+1)$.
Since we start at a vertex, $\E[X_i]=0$.
Then
\begin{align}
\tcov&=   0+\sum_{i=2}^{n} \frac{n-1}{n-i+1} \nonumber \\
&=  (n-1) \left(\frac{1}{n-1}+\frac{1}{n-2}+\cdots+1\right). \nonumber
\end{align}

A natural interpretation of the cover time on a complete graph is 
the coupon collector problem we introduced in \cref{sec:intro}.
There are $r$ distinct coupons one of which arrives in the mail everyday.
We want to know how many days until the collector has collected each coupon.
The strategy is to think of every coupon as a vertex on the complete graph.
We then move from vertex to vertex with uniform probability.
The difference between the coupon collector problem and the complete graph
is that we now have self-edges
whereas before we had to leave our current vertex at each step.
That is, we can now stay in the same vertex (provided the collector 
receives the same coupon two days in a row).
We apply our approach to the complete graph and substitute the $r$ possible coupons
we can find for the $n-1$ vertices we could move to.
Then
\begin{align}
\tcov&=  r \left(\frac{1}{r-1}+\frac{1}{r-2}+\cdots+1\right). \nonumber
\end{align}

@@

We will begin by considering the time a random walk takes to
visit either one of the endpoints from vertex $i$.
From $X_t=i$, we move with equal probability to $i+1$
and $i-1$ until we reach either $0$ or $N$.

@@

where $\E_{i}[T_{i+1}]$ is the expected time from state $i$ to $i+1$.
Since there is only one edge incident to 0, we always move from 0 to 1.
From state $i\neq0$, we take one step in each direction with equal probability.
In half the cases, we have arrived at $i+1$.
In the other half of cases, we are in $i-1$.
It will take us $\E_{i-1}[T_i]$ steps back to $i$ and then another
$\E_i[T_{i+1}]$ steps to $i+1$.
Formally,
\begin{align}
\E_{i}[T_{i+1}] &= 1 + \frac{1}{2}(\E_{i-1}[T_i]+\E_i[T_{i+1}]) +\frac{1}{2}0 \nonumber \\
 &= 2 + \E_{i-1}[T_{i}]. \nonumber
\end{align}

@@

Then
\begin{align}
\E_0[T_N] &= 1 + 3 + \cdots + 2i - 1 + \cdots + 2(N-1) -1 = N^2 \nonumber
\end{align}

By \cref{thm:line}, for the random walk from state $i$ on the line graph,
\begin{align}
\tcov = i(N-i) + N^2. \nonumber
\end{align}

@@

Consider a random walk on a simple cycle as in \cref{fig:cycle}.
(Note that the starting vertex is arbitrary since the graph is symmetric.)
Let $(X_t)$ be the walk on the number line (between $-n$ and $n$)
where $X_0 = 0$.
The cover time $\tcov$ of the simple cycle is the expected number of steps 
until we visit $n$ distinct integers on the number line.

@@

\section{Cover Time Bounds}\label{sec:bounds}

@@

Recall $\sum_y P_{a,y} v_y$ is the probability we reach $a$ before $b$
so the complement is the probability we reach $b$ before returning to $a$.

@@

Recall that $\pi_a$ is the stationary probability that $(X_t) = a$.
By a similar argument, $1/\pi_a$ is the expected return time to $a$.
If we condition on reaching $b$ before turning to $a$,
\begin{align}
\frac{1}{\pi_a} &= \E_a[T_a] =
\E_a[T_a|T_b < T_a] \pr(T_b<T_a) +
\E_a[T_a|T_a < T_b] \pr(T_a<T_b) \nonumber.
\end{align}

The expected commute time between $a$ and $b$ is the
random sum expectation,
\begin{align}
\E_a[T_b] + \E_b[T_a] &= 
\E_a[T_a | T_b < T_a] + 
\E_a[T_a | T_a < T_b] (\frac{1}{\pr(T_b<T_a)} - 1) \nonumber.
\end{align}
Observe that $1/\pr(T_b<T_a)-1 = (1-\pr(T_b<T_a))/\pr(T_b<T_a)$, 
which simplifies to $\pr(T_a<T_b)/\pr(T_b<T_a)$.
Multiplying by the probability we reach $b$ before returning to $a$,
\begin{align}
\E_a[T_a|T_b < T_a] \pr(T_b<T_a) + \E_a[T_a|T_a<T_b] \pr(T_a<T_b)=\frac{1}{\pi_a} \nonumber
\end{align}
and the commute identity follows
\begin{align}
E_a[T_b] + E_b[T_a] = \frac{1}{\pi_a \pr_a(T_b < T_a)} = \Reff C.
\label{eqn:commute}
\end{align}

@@

Our goal is to bound the cover time in terms of the hitting time.
Consider the first term of the sum.
Recall that the hitting time $\hit$ is the maximum
expected time from any vertex to any other.
It follows that for any choice of $\sigma_1$,
the expected time from $n$ to $\sigma_1$ is less than or equal to $\hit$.
Formally for $s$ an arbitrary state such that $1\leq s \leq n-1$,
$\E_n[T_1] = \E_n[T_s] \leq \hit$.

@@

Since $\sigma$ is a random permutation, the probability that
we visit $\sigma_k$ last is $1/k$.
Now we find the expected time between $T_{k-1}$ and $T_{k}$
given that last vertex we visit is $\sigma_k$.
Consider $r$ and $s$ where $1 \leq r \neq s \leq n-1$
such that $L_{k-1} = r$ and $L_k = s$.
Then $\E_n[T_k - T_{k-1} | L(k) = \sigma_k] = \E_r[T_s] \leq \hit$.
So
\begin{align}
\E_n[T_k - T_{k-1}] &\leq
\hit \frac{1}{k}. \nonumber 
\end{align}

@@

The same strategy for the Matthews upper bound can be used
to find a lower bound.
Instead of looking for all $n$ vertices, we search
for $A \subseteq V(G)$.
The Matthews lower bound will always hold.
However, we can tighten the bound by picking $A$ such
that the hitting time is large between any pair of vertices in $A$.
We use `large' in a relative sense and 
offer an illustrative example in \cref{sec:binary}.

@@

\begin{proof}
Without loss of generality, label the vertices $\{1, 2, \ldots, x\}$
such that $x=|A|$ and vertex $x$ is the vertex from which it takes the
most steps to cover the other vertices.
Let $\sigma$ be a uniform permutation of the vertices in $A$.
The strategy is to look for the vertices in order $\sigma$.

@@

Our goal is to bound the cover time in terms of $\tmina$.
Consider the first term of the sum.
With probability $1/|A|$, $\sigma_1 = x$ and $T_1=0$.
Otherwise, the expected time from $x$ to $T_1$ is less than or equal to
$\tmina$.
Then
\begin{align}
\E_x[T_1] \geq \frac{1}{|A|}0 + \frac{|A| -1}{|A|} \tmina
= \left(1 - \frac{1}{|A|} \right) \tmina \nonumber
\end{align}

@@

Since $\sigma$ is a random permutation, $P(L_k = \sigma_k) = 1/k$.
Now find the expected time between $T_{k-1}$ and $T_{k}$
given that last vertex we visit is $\sigma_k$.
Consider $r, s \in A$ such that $L_{k-1} = r$ and $L_k = s$.
Then $\E_x[T_k - T_{k-1} | L(k) = \sigma_k] = \E_r[T_s] \geq \tmina$.
So
\begin{align}
\E_x[T_k - T_{k-1}] &\geq
\tmina \frac{1}{k}. \nonumber 
\end{align}

@@

We now use the Matthews Lower Bound.
We want to choose a set $A \subseteq V(G)$ with a large hitting time between any
pair of vertices.
To find such a set, fix a level $h$ in our binary tree.
Then let $A$ be a set of $2^h$ leaves so that each vertex at level $h$
has a unique descendent in $A$.
As $h$ increases, the distance between the leaves in $A$ decreases.
Thus we want to choose a smaller value of $h$ so that our lower bound will be bigger
(and better).

@@

A spanning tree $\T$ is a subset of a graph with all vertices 
in $V(G)$ connected by the minimum number of edges in $E(G)$.
\cref{fig:span} is an example where the bold edges indicate the edges in $E(\T)$
and the light edges indicate the edges in $E(G)$ but not $E(\T)$.
For the connected graphs we consider, the minimum number of edges
needed to connect all vertices is $n-1$.
Formally, $V(\T) = V(G)$ and $E(\T) \subseteq E(G)$
such that $|E(\T)| = n-1$ and $\T$ is connected.

@@

The Poisson approximation of the binomial implies that if $n \rightarrow \infty$
and $r/n \rightarrow c$, then the number of visits to each vertex will approach
a Poisson distribution with mean $c$.
We can use this to intuite our latest result: the proportion of visits to each vertex
approaches $e^{-c}$.

@@

Define $p_m(r,n)$ as the probability that exactly $m$ of $n$ vertices are unvisited
when $r$ steps have been taken.
Then the probability that all vertices have been visited is the complement
of the probability that at least one vertex has not been visited.

@@

We bound the expression from the top and the bottom and use the squeeze
theorem to prove the most recent limit.
We will first show the upper bound.
Notice that $(1-x) \leq e^{-x}$ and recall our assumption that
$ne^{r/n} \rightarrow \lambda$.
Then
\begin{align}
{n \choose m} \left(1-\frac{m}{n}\right)^r &\leq \frac{n^m}{m!} e^{mr/n} 
= \frac{(n e^{r/n})^m} {m!}
\rightarrow \frac{\lambda ^m} {m!}. \label{eqn:up_lim}
\end{align}

@@

\subsection{Birthday Problem}\label{sec:birthday}
Consider a complete graph (with self-edges) and 365 vertices.
We may think of the cover time variable of this graph as the number of people
until every birthday is represented.
We call this the birthday problem.

We use \cref{eqn:dist} to calculate the probability
that all birthdays are represented in populations of 
size $r=1825$ and $r=2190$, respectively.
The number of vertices $n$ is 365 so $n\log n = 2153$.
It follows that 
\begin{align}
\pr(\cov \leq 1825) &= \pr \left(\frac{\cov - 2153}{365} \leq
\frac{-328}{365}\right) \nonumber \\
&\approx e^{-e^{-0.89863}} \approx 0.08576 \nonumber
\end{align}
but
\begin{align}
\pr(\cov \leq 2190) &= \pr \left(\frac{\cov - 2153}{365} \leq
\frac{37}{365}\right) \nonumber \\
&\approx e^{-e^{-0.10137}} \approx 0.40511. \nonumber
\end{align}
